# Week 3 exercises

## Data

```{r}
library(readr)
alc <- read_csv("data/alc.csv")
dim(alc)
str(alc)
colnames(alc)
# ... the variables are as listed by the above function, with gender being of chr type, rest numeric.
```
The data in 'alc' describes questionnaire data on students in two Portuguese schools. It has student metadata, and details on their school performance and alcohol consumption.

The purpose of your analysis is to study the relationships between high/low alcohol consumption and some of the other variables in the data. To do this, choose 4 interesting variables in the data and for each of them, present your personal hypothesis about their relationships with alcohol consumption. (0-1 point)

## Hypothesis

1. High alcohol usage probably correlates with number of absences.
2. Volume of alcohol usage is likely strongly gendered.
3. High alcohol usage probably weakly inversely correlates with 'G3' (final year grade).
4. High alcohol usage likely correlates with going out with friends ('goout').

## Data exploration

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
alc_subset = select(alc, c('alc_use', 'high_use', 'sex', 'goout', 'G3', 'absences'))

gather(alc_subset) %>% glimpse()

gather(alc_subset) %>% View()

gather(alc_subset) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
```

Absences and G3 score seem to have high numbers in both ends of the scale. Absences also has a strange peak in the middle, which might be an artifact of school practices. Maybe X amount of absences either automatically disqualifies one from the class, or X amount is the maximum permitted amount. Alcohol use is weighted towards the lower end of the scale, out-goingness is somewhat normally distributed and male:female ratio is close to 1:1.

```{r}
library(dplyr)
alc_subset %>% group_by(sex, high_use) %>% summarise(count = n())
```

Males are somewhat over represented in the high alcohol usage group, as hypothesised.

```{r}
alc_subset %>% group_by(sex, high_use) %>% summarise(count = n(), mean_grade = mean(G3))

```

High alcohol use seems to somewhat lower the grades in males, but not in females.

```{r}
# Some of the exercise code fit with my hypothesis nicely...
g1 <- ggplot(alc, aes(x = high_use, y = G3))
g1 + geom_boxplot(aes(col = sex)) + ylab("grade")
```

And the same can be seen from the above plot.

```{r}
alc_subset %>% group_by(high_use) %>% summarise(mean_goout = mean(goout))

```
Going out seems to quite clearly correlate with high alcohol usage, as was the hypothesis.

```{r}
alc_subset %>% group_by(high_use) %>% summarise(mean_absences = mean(absences))
```

And alcohol use and absences correlate.

```{r}
g2 <- ggplot(alc, aes(x = high_use, y = absences))
g2 + geom_boxplot(aes(col = sex)) + ylab("absences") +
  ggtitle("Student absences by alcohol consumption and sex")
```
But gender doesn't seem to make a huge difference here. Still, alhocol use in males does correlate with absences more clearly.

## Logistical regression models

```{r}
m <- glm(high_use ~ absences + goout + sex + G3, data = alc_subset, family = "binomial")
summary(m)
```
absences, goout, and gender all strongly correlate with high_use, as demonstrated by the low p-scores, but G3 clearly does not. So, let's drop G3 from the model:

```{r}
m <- glm(high_use ~ absences + goout + sex, data = alc_subset, family = "binomial")
summary(m)
```
And add coefficients and confidence intervals:

```{r}
OR <- coef(m) %>% exp
CI <- confint(m) %>% exp()
cbind(OR, CI)
```
Alcohol consumption seem to here have a significantly lower effect on on absences, than how the results earlier seemed. Apparently both male gender and goout would already serve to explain majority of the effect covered by absences?

## Cross tabulation

Using the variables which, according to your logistic regression model, had a statistical relationship with high/low alcohol consumption, explore the predictive power of you model. Provide a 2x2 cross tabulation of predictions versus the actual values and optionally display a graphic visualizing both the actual values and the predictions. Compute the total proportion of inaccurately classified individuals (= the training error) and comment on all the results. Compare the performance of the model with performance achieved by some simple guessing strategy. (0-3 points)

Let's see if I have time for this ...
