# Week 3 exercises

## Data

```{r}
library(readr)
alc <- read_csv("data/alc.csv")
dim(alc)
str(alc)
colnames(alc)
# ... the variables are as listed by the above function, with gender being of chr type, rest numeric.
```
The data in 'alc' describes questionnaire data on students in two Portuguese schools. It has student metadata, and details on their school performance and alcohol consumption.

The purpose of your analysis is to study the relationships between high/low alcohol consumption and some of the other variables in the data. To do this, choose 4 interesting variables in the data and for each of them, present your personal hypothesis about their relationships with alcohol consumption. (0-1 point)

## Hypothesis

1. High alcohol usage probably correlates with number of absences.
2. Volume of alcohol usage is likely strongly gendered.
3. High alcohol usage probably weakly inversely correlates with 'G3' (final year grade).
4. High alcohol usage likely correlates with going out with friends ('goout').

## Data exploration

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
alc_subset = select(alc, c('alc_use', 'high_use', 'sex', 'goout', 'G3', 'absences'))

gather(alc_subset) %>% glimpse()

gather(alc_subset) %>% View()

gather(alc_subset) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
```

Absences and G3 score seem to have high numbers in both ends of the scale. Absences also has a strange peak in the middle, which might be an artifact of school practices. Maybe X amount of absences either automatically disqualifies one from the class, or X amount is the maximum permitted amount. Alcohol use is weighted towards the lower end of the scale, out-goingness is somewhat normally distributed and male:female ratio is close to 1:1.

```{r}
library(dplyr)
alc_subset %>% group_by(sex, high_use) %>% summarise(count = n())
```

Males are somewhat over represented in the high alcohol usage group, as hypothesised.

```{r}
alc_subset %>% group_by(sex, high_use) %>% summarise(count = n(), mean_grade = mean(G3))

```

High alcohol use seems to somewhat lower the grades in males, but not in females.

```{r}
# Some of the exercise code fit with my hypothesis nicely...
g1 <- ggplot(alc_subset, aes(x = high_use, y = G3))
g1 + geom_boxplot(aes(col = sex)) + ylab("grade")
```

And the same can be seen from the above plot.

```{r}
alc_subset %>% group_by(high_use) %>% summarise(mean_goout = mean(goout))

```
Going out seems to quite clearly correlate with high alcohol usage, as was the hypothesis.

```{r}
alc_subset %>% group_by(high_use) %>% summarise(mean_absences = mean(absences))
```

And alcohol use and absences correlate.

```{r}
g2 <- ggplot(alc_subset, aes(x = high_use, y = absences))
g2 + geom_boxplot(aes(col = sex)) + ylab("absences") +
  ggtitle("Student absences by alcohol consumption and sex")
```
But gender doesn't seem to make a huge difference here. Still, alhocol use in males does correlate with absences more clearly.

## Logistical regression models

```{r}
m <- glm(high_use ~ absences + goout + sex + G3, data = alc_subset, family = "binomial")
summary(m)
```
absences, goout, and gender all strongly correlate with high_use, as demonstrated by the low p-scores, but G3 clearly does not. So, let's drop G3 from the model:

```{r}
m <- glm(high_use ~ absences + goout + sex, data = alc_subset, family = "binomial")
summary(m)
```
And add coefficients and confidence intervals:

```{r}
OR <- coef(m) %>% exp
CI <- confint(m) %>% exp()
cbind(OR, CI)
```
Alcohol consumption seem to here have a significantly lower effect on on absences, than how the results earlier seemed. But the scale for absences is a lot more granular than for the other variables here, so that might have an effect? Or maybe the effect isn't as pronounced as on the other variables. There's roughly 50% chance (1:1 odds) that there's a unit change in absences if alcohol consumption state flips from high to low. For gender that chance is ~75% (2.78:1) and ~65% chance for out-goingness.

## Cross tabulation of predictions

```{r}
probabilities <- predict(m, type = "response")
alc_subset <- mutate(alc_subset, probability = probabilities)
alc_subset <- mutate(alc_subset, prediction = probability > 0.5)
raw_num <- table(high_use = alc_subset$high_use, prediction = alc_subset$prediction)
raw_num
table(high_use = alc_subset$high_use, prediction = alc_subset$prediction) %>%
  prop.table() %>%
  addmargins()

precision <- raw_num[2,2]/sum(raw_num[,2])
precision

recall <- raw_num[2,2]/sum(raw_num[2,])
recall
```
The model doesn't seem that great. Precision was *~75%*, that is only 3/4 of the positive predictions were accurate. Recall is *~45%*, so more than half of actual high_use cases were undetected.

```{r}
1 - (sum(alc_subset$high_use == alc_subset$prediction) / nrow(alc_subset))
```
Total proportion of inaccurate classifications is *~21%*. That's not too great a result. Seems like alcohol consumption does not explain all other behaviour, but still a surprisingly high number of predictions are correct. The relatively good result is probably because the model seems to err on the side of not high use, and as that is the more likely result the number of accurate predictions is misleadingly high. Precision and recall, (which were not great) are probably better indicators.

If we try predicting with the three variables we decided were significant, by predicting high alcohol consumption if at least two of the three are on the side of the indicator favoring high alcohol consumption, the results look like this:

```{r}
alc_subset$goout_guess <- alc_subset$goout > mean(alc_subset$goout)
alc_subset$abs_guess <- alc_subset$absences > mean(alc_subset$absences)
alc_subset$gender_guess <- alc_subset$sex == "M"
alc_subset$ultimate_guess <- (alc_subset$abs_guess + alc_subset$goout_guess + alc_subset$gender_guess) >= 2
1 - (sum(alc_subset$high_use == alc_subset$ultimate_guess) / nrow(alc_subset))
```
We got 28% wrong, which is a bit higher than the model's 21%, but not that high. I wouldn't base students' alcoholism intervention program on either.
