# Week 6 exercises

Couldn't finish these in time. Apologies to whoever is checking this.

```{r}
RATS <- read.table(file = "data/rats.csv", sep = ",", header = TRUE)
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)

BPRS <- read.table(file = "data/bprs.csv", sep = ",", header = TRUE)
BPRS$treatment <- factor(BPRS$treatment) 
BPRS$subject <- factor(BPRS$subject) 
```

## Overview of RATS

```{r}
library(ggplot2)

# Draw the plot
ggplot(RATS, aes(x = Time, y = weight, color = ID, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATS$weight), max(RATS$weight)))
```
Might have made more sense to have them all in one graph, with linetype = Group instead of ID.


```{r}
library(dplyr)
library(tidyr)

RATS <- RATS %>%
  group_by(Time) %>%
  mutate(stdweight = (weight - mean(weight))/ sd(weight)) %>%
  ungroup()

# Plot again with the standardized
ggplot(RATS, aes(x = Time, y = stdweight, color = ID, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:16, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")

```
Standardized version.

## Summary of rat weights per week

```{r}
n <- 16

# Summary data with mean and standard error of weight by treatment and week 
RATSS <- RATS %>%
  group_by(Group, Time) %>%
  summarise(mean = mean(weight), se = sd(weight)/n ) %>%
  ungroup()

# Plot the mean profiles
library(ggplot2)
ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.9,0.5)) +
  scale_y_continuous(name = "mean(weight) +/- se(weight)")

```

And mean weights:

```{r}
RATS8S <- RATS %>%
  group_by(Group, ID) %>%
  summarise(mean=mean(weight)) %>%
  ungroup()

library(ggplot2)
ggplot(RATS8S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(weight)")
```

Had to filter outliers from each Group of rats separately.

```{r}
RATS8S_filtered <- RATS8S %>% filter(!(Group == 1 & mean < 250) &
                                     !(Group == 2 & mean > 550) &
                                     !(Group == 3 & mean < 500))

ggplot(RATS8S_filtered, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(bprs), weeks 1-8")
```

... and the groupings are very tight as a results of filtering. I wonder if it did really make sense?

## T-test, Anova

t-test only makes sense pairwise, so we'll filter for groups. Could make a t-test of group 1 vs group 3 too, but that'd show even more obvious significance than these tests. 

```{r}
RATS12 <- RATS8S_filtered %>% filter(Group == 1 | Group == 2)
t.test(mean ~ Group, data = RATS12, var.equal = TRUE)

RATS23 <- RATS8S_filtered %>% filter(Group == 2 | Group == 3)
t.test(mean ~ Group, data = RATS23, var.equal = TRUE)

# Add the baseline from the original data as a new variable to the summary data
RATS_bl <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", header = TRUE, sep = '\t')


RATS8S_non_filtered <- RATS8S %>% mutate(baseline = RATS_bl$WD1)

# Fit the linear model with the mean as the response 
fit <- lm(mean ~ baseline + Group, data = RATS8S_non_filtered)

# Compute the analysis of variance table for the fitted model with anova()
anova(fit)

```
Looks like Group matters, as was evident from the graphs too... but it was practically all determined by the baseline. The p-value of 0.07 for 'Group' barely registers as signifying at all.

